---
title: "Data Science Capstone"
author: "Shun-Wen Chang"
date: "September 25, 2016"
output: html_document
---

This is the milestone report for Data Science Capstone project from Coursera Data Science Specialization. The objectives of this report is to load the 3 given data sets, summarize the data, and explore the data to understand the frequency distribution of words and 2-gram, 3-gram words.

###0. Examine Data###

Before loading data, let's check the data size and word counts within bash shell.

**check file sizes**<br>
-rw-r--r--@ 1 EstherChang  staff  167105338 Jul 22  2014 en_US.twitter.txt<br>
-rw-r--r--@ 1 EstherChang  staff  205811889 Jul 22  2014 en_US.news.txt<br>
-rw-r--r--@ 1 EstherChang  staff  210160014 Jul 22  2014 en_US.blogs.txt<br>

**check line counts**<br>
  899288 en_US.blogs.txt<br>
 1010242 en_US.news.txt<br>
 2360148 en_US.twitter.txt<br>
 4269678 total<br>


###1. Load Data and necessary packages###

```{r, echo=TRUE}
setwd("/Users/EstherChang/Documents/R/Data_Science_captone/final/en_US/")
blog <- readLines("en_US.blogs.txt",skipNul = TRUE, warn = TRUE)
news <- readLines("en_US.news.txt",skipNul = TRUE, warn = TRUE)
twitter <- readLines("en_US.twitter.txt",skipNul = TRUE, warn = TRUE)

library(ggplot2)
library(NLP)
library(tm)
library(RWeka)

```


###2. Data Sampling###
Because these data are huge. We need to make a sample subset to do this project. In each file, I select random 1000 entries as my data source and then delete the original data to release memory space.

```{r, echo=TRUE}
set.seed(100)
sample_size = 1000

sample_blog <- blog[sample(1:length(blog),sample_size)]
sample_news <- news[sample(1:length(news),sample_size)]
sample_twitter <- twitter[sample(1:length(twitter),sample_size)]

```

**Examing the first few lines of each data set:**
```{r, echo=TRUE}
head(sample_blog)
head(sample_twitter)
head(sample_news)
```

**Then combine all 3 data and remove originals:**

```{r, echo=TRUE}

sample_data<-rbind(sample_blog,sample_news,sample_twitter)
rm(blog,news,twitter)

```

###3. Clean Data###

**I clean the data with following rules:**
1. remove punctuation
2. remove whitespace
3. discard numbers since they are irrelavant in our analysis
4. convert to all lowercases


**Clean the data using tm_map:**

```{r, echo=TRUE}
mycorpus<-VCorpus(VectorSource(sample_data))
mycorpus <- tm_map(mycorpus, content_transformer(tolower)) # convert to lowercase
mycorpus <- tm_map(mycorpus, removePunctuation) # remove punctuation
mycorpus <- tm_map(mycorpus, removeNumbers) # remove numbers
mycorpus <- tm_map(mycorpus, stripWhitespace) # remove multiple whitespace
changetospace <- content_transformer(function(x, pattern) gsub(pattern, " ", x))
mycorpus <- tm_map(mycorpus, changetospace, "/|@|\\|")

```

###4. Tokenize the sentences###

We use NGramTokenizer in RWeka package for this task. In this project, we analyze 1gram, 2gram, and 3gram, which I will call "oneGM", "twoGM", and "threeGM", respectively for the n-gram matrices.



```{r, echo=TRUE}

uniGramTokenizer <- function(x) NGramTokenizer(x, Weka_control(min = 1, max = 1))
biGramTokenizer <- function(x) NGramTokenizer(x, Weka_control(min = 2, max = 2))
triGramTokenizer <- function(x) NGramTokenizer(x, Weka_control(min = 3, max = 3))
OneT <- NGramTokenizer(mycorpus, Weka_control(min = 1, max = 1))
oneGM <- TermDocumentMatrix(mycorpus, control = list(tokenize = uniGramTokenizer))
twoGM <- TermDocumentMatrix(mycorpus, control = list(tokenize = biGramTokenizer))
threeGM <- TermDocumentMatrix(mycorpus, control = list(tokenize = triGramTokenizer))

```

###5. Generate n-gram histograms###

**Unigram frequency**

```{r, echo=TRUE}
freqTerms <- findFreqTerms(oneGM, lowfreq = 200)
termFreq <- rowSums(as.matrix(oneGM[freqTerms,]))
termFreq <- data.frame(unigram=names(termFreq), frequency=termFreq)

g1 <- ggplot(termFreq, aes(x=reorder(unigram, frequency), y=frequency)) +
    geom_bar(stat = "identity") +  coord_flip() +
    theme(legend.title=element_blank()) +
    xlab("Unigram") + ylab("Frequency") +
    labs(title = "Top unigrams by frequency")
print(g1)

```

**Bigram frequency**

```{r, echo=TRUE}
freqTerms <- findFreqTerms(twoGM, lowfreq = 70)
termFreq <- rowSums(as.matrix(twoGM[freqTerms,]))
termFreq <- data.frame(bigram=names(termFreq), frequency=termFreq)

g2 <- ggplot(termFreq, aes(x=reorder(bigram, frequency), y=frequency)) +
    geom_bar(stat = "identity") +  coord_flip() +
    theme(legend.title=element_blank()) +
    xlab("Bigram") + ylab("Frequency") +
    labs(title = "Top bigrams by frequency")
print(g2)

```

**Trigram frequency**

```{r, echo=TRUE}
freqTerms <- findFreqTerms(threeGM, lowfreq = 10)
termFreq <- rowSums(as.matrix(threeGM[freqTerms,]))
termFreq <- data.frame(trigram=names(termFreq), frequency=termFreq)

g3 <- ggplot(termFreq, aes(x=reorder(trigram, frequency), y=frequency)) +
    geom_bar(stat = "identity") +  coord_flip() +
    theme(legend.title=element_blank()) +
    xlab("Trigram") + ylab("Frequency") +
    labs(title = "Top trigrams by frequency")
print(g3)

```
